---
title: "HiTMapper"
author: "Matei"
date: "2/24/2021"
output: 
  html_document:
    fig_width: 12
    fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HiTMapper)
library(ggraph)
library(ggiraph)
library(magrittr)
```

## R Markdown

Load the example dataset, originally from the FlowSOM package.

```{r lymphocytes}
data("lymphocytes")
head(lymphocytes)
```


Obtain the network and prune some low-weight edges (to speed up the plotting 
without noticeable impact on performance). 
Use the Leiden method to detect communities.

```{r mapper}

set.seed(52)
mapper <- HiTMapper(lymphocytes, kNodes = 400, overlap = 0.3,
                    outlierCutoff = 25)
mapper <- pruneEdges(mapper, cutoff=0.05)

community <- leidenClustering(mapper$gr, resolution = 2.5)
```

Create a 2D layout of the network and save it for downstream plotting.

```{r layout}
layout <- create_layout(mapper$gr, layout="fr")
size <- sapply(mapper$nodes, length)
```

Label the communities, using human knowledge about the markers.

```{r label}
mapping <- assignCells(lymphocytes, mapper, community)

medians <- sapply(levels(community), function(l) {
  apply(lymphocytes[which(mapping==l),], 2, median)
}) %>% t()

binary <- medians > 1.7

labels <- rep("Myeloid", nrow(medians))
labels[which(binary[,"CD4"])] <- "T cell CD4"
labels[which(binary[,"CD8"])] <- "T cell CD8"
labels[which(binary[,"NK1.1"])] <- "NK cell"
labels[which(binary[,"CD19"])] <- "B cell"
labels[which(binary[,"TCRyd"])] <- "T cell gd"

labels[which(!binary[,"CD45"])] <- "Not leukocyte"
labels[which(binary[,"NK1.1"]&binary[,"CD3"])] <- "NK T cell"
labels[which(binary[,"CD8"]&!(binary[,"CD3"]))] <- "CD8 no CD3"
labels[which(binary[,"TCRb"]&!(binary[,"CD8"])&
            !(binary[,"CD4"])&!(binary[,"NK1.1"]))] <- "T cell other"
cellType <- as.factor(labels[community])
```

Plot the cell type assignment.

```{r communities}
ggraph(layout) +
  geom_edge_link(aes(alpha = weight)) +
  geom_node_point(aes(color = cellType, size=size)) +
  scale_color_brewer(palette = "Paired") +
  scale_edge_alpha(guide="none") +
  theme_graph(base_family = "sans") +
  theme(text=element_text(size = 18))
```


Check the distribution of various markers in the Mapper graph, 
and their agreement with the cell types.

```{r distribution}
markers <- c("CD3", "CD8", "CD19", "NK1.1")
for(marker in markers) {
  color <- mapper$nodeStats$q50[,marker]
  g <- ggraph(layout) +
    geom_edge_link(aes(alpha = weight)) +
    geom_node_point(aes(color=color, size=size)) +
    scale_color_gradient(low="black", high="red", name=marker) +
    scale_edge_alpha(guide="none") +
    theme_graph(base_family = "sans") +
    theme(text=element_text(size = 18))

  plot(g)
}

```


