---
title: "HiTMapper"
author: "Matei"
date: "2/24/2021"
output: 
  html_document:
    fig_width: 12
    fig_height: 10
---

## First install and load the relevant packages

* HiTMapper
* ggraph
* tidyverse

## Import the packages we need

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HiTMapper)
library(ggraph)
library(tidyverse)
```

## Load the example dataset from the HiTMapper package.
When HiTMapper is imported, an example data set is also loaded. Options to load the example data include running any of the following: HiTMapper::toy_data, toy_data, or data("toy_data").

```{r toy_data}
data("toy_data")
head(toy_data)
```


Obtain the network and prune some low-weight edges (to speed up the plotting 
without noticeable impact on performance). For the moment, you need to provide
the data in matrix or data frame format, aggregated from all fcs files of interest.

```{r mapper}
set.seed(52)
mapper <- HiTMapper(toy_data, total_nodes=200)
#mapper <- pruneEdges(mapper, cutoff=0.01)
```

The function `getCommunitiesFeatures()` is a wrapper around three functionalities:
1. community detection using the Leiden algorithm;
2. computing median marker expression for each community, and labeling communities based on a table of user-provided phenotypes;
3. computing a sample-by-community contingency table (the "features" are percentages of cells from each sample which belong to a given community).
It needs as input the phenotype definitions and the assignment of cells to biological samples.

Load the phenotype table. Phenotypes should be listed along the rows, with "lo", "hi" used to indicate
the marker modality which defines the phenotype. Missing values, or values other than "lo", "hi", indicate
markers which do not contribute to the phenotype definition.
```{r defs}
#data("defs")
#defs
```

In the example data, we have a single sample.
```{r samples}
#sample_mapping <- rep("1", nrow(toy_data))
```

Compute communities and features, and augment the mapper object with the result. 
The `resolution` parameter is passed to the Leiden method;
a higher resolution means more resulting communities.
```{r communities_features}
#mapper <- getCommunitiesFeatures(toy_data, sample_mapping, mapper, defs, resolution=2)
#levels(mapper$community)
#mapper$features
```

You can use the `ggraph` package to make various network plots;
just provide the `gr` field of the mapper object, which contains
the graph. For example, color the nodes by community assignment.
The layout is non-deterministic; try out a couple of random seeds.

```{r communities}
#set.seed(52)
#community <- mapper$community
#size <- sapply(mapper$nodes, length)

#ggraph(mapper$gr, layout="fr") +
#  geom_edge_link(aes(alpha = weight)) +
#  geom_node_point(aes(color = community, size=size)) +
#  scale_edge_alpha(guide="none") +
#  theme_graph(base_family = "sans") +
#  theme(text=element_text(size = 18))
```

Alternatively, use the built-in `plotMapper` function, which
plots the distribution of various markers in the Mapper graph, 
and their agreement with the community assignment. The `path`
argument can be used to save plots to files; this is recommended
when plotting multiple markers.

```{r distribution}
set.seed(52)
markers <- c("CD3", "CD8", "CD19", "NK1.1")
plot_mapper(mapper, markers=markers, path=NULL)
```

Also check out the function `plotMapperInteractive`, which outputs
interactive network plots in html format.




