---
title: "HiTMapper"
author: "Matei"
date: "2/24/2021"
output: 
  html_document:
    fig_width: 12
    fig_height: 10
---

## First install and load the relevant packages

* HiTMapper
* tidyverse

## Import the packages we need

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HiTMapper)
library(ggraph)
library(tidyverse)
```

## Load the example dataset from the HiTMapper package.
When HiTMapper is imported, an example data set is also loaded. Options to load the example data include running any of the following: HiTMapper::toy_data, toy_data, or data("toy_data").

```{r toy_data}
data("toy_data")
head(toy_data)
```

A data frame containing phenotype definitions is also available in the package.
HiTMapper can use it to label clusters. For different datasets/panels, you can
provide phenotype definitions in the same format.

```{r defs}
defs
```


The function `HiTMapper` is a wrapper for all important functionality.
The only required parameter is the input cytometry data.

* For the moment, you need to provide the data as a matrix, 
aggregated from all fcs files of interest.
* The default number of nodes to be used in the graph is 1000; 
for small datasets like our example data (10,000 cells) a smaller number 
works better.
* After building the graph, HiTMapper groups similar nodes into larger
communities. The resolution parameter controls the granularity of communities.
Values between 0.1 and 2 are usually good; the default is 1.

The output of `HiTMapper` is a list, containing various 
information about the model.

```{r mapper}
set.seed(0)
mapper <- HiTMapper(toy_data, total_nodes=200, resolution=0.5, defs=defs)
names(mapper)
```


For the remainder of this vignette, we will explore some of the list elements
in the output. The most important is `clustering`, an array of cluster membership
ids for every cell. You can use the `table` function from base R to get a quick
summary.

```{r clustering}
table(mapper$clustering)
```

Another important list element is `community_medians`, which gives median
marker intensity for each community. Internally, `HiTMapper` uses these
community medians and the phenotype definitions to assign labels.

```{r community_medians}
heatmap(mapper$community_medians)
```

The list element `gr` contains the graph that HiTMapper constructed,
as an igraph object. If you're familiar with the `igraph` package, you
can use the functions therein to operate on `gr`.
You can also use the `ggraph` package to make various network plots;
for example, color the nodes by community assignment.
The layout is non-deterministic; try out a couple of random seeds.

```{r communities}
set.seed(0)

ggraph(mapper$gr, layout="fr") +
 geom_edge_link(aes(alpha = weight)) +
 geom_node_point(aes(color = mapper$community, size=tabulate(mapper$mapping))) +
 scale_edge_alpha(guide="none") +
 theme_graph(base_family = "sans") +
 theme(text=element_text(size = 18))
```

Alternatively, use the built-in `plot_mapper` function, which
plots the distribution of chosen markers in the Mapper graph, 
as well as the community assignment of each node. The `path`
argument can be used to save plots to files; this is recommended
when plotting multiple markers.

```{r distribution}
set.seed(0)
markers <- c("CD3", "CD19", "CD16")
plot_mapper(mapper, markers=markers, path=NULL)
```





